---
title: "Cyclistic Analysis"
author: "Airel Camilo"
date: "2023-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Process

Install the packages
```{r}
install.packages("tidyverse")
library(tidyverse)
library(dplyr)
library(readr)
```

Merge the six dataset into one called tripdata, remove null value and duplicate rows
```{r}
tripdata <- list.files(path="divvy-tripdata", full.names = TRUE) %>%
  lapply(read_csv) %>% 
  bind_rows %>%
  drop_na()

tripdata <- tripdata[!duplicated(tripdata$ride_id), ]
```

Add columns such as, ride_length, hour_used, day_used, and month_used
```{r}
tripdata_new <- tripdata %>%
  mutate(ride_length = difftime(ended_at, started_at, units="mins")) %>%
  mutate(hour_used = format(started_at, "%H")) %>%
  mutate(day_used = format(started_at, "%A")) %>%
  mutate(month_used = format(started_at, "%B"))
```

Drop unnecessary fields, remove negative ride_length and outlier in ride_length
```{r}
tripdata_final <- subset(tripdata_new, select = -c(started_at,ended_at,start_station_name,start_station_id, end_station_name,end_station_id,start_lat,start_lng,end_lat,end_lng))

tripdata_final <- subset(tripdata_final, ride_length > 0)

Q <- quantile(tripdata_final$ride_length, probs=c(.25,.75)) # Find Q1 and Q3 for calculating RUB
# RUB = Q3 + 1.5(Q3-Q1)
RUB <- Q[2]  + 1.5*(Q[2]-Q[1])
tripdata_final <- subset(tripdata_final, ride_length < RUB)

```

Make different table for each categories and group data by member type
```{r}
member_per_group <- tripdata_final %>%
  group_by(member_casual) %>%
  summarize(total_ride = n(), avg_ride_length = mean(ride_length))

bike_type_per_group <- count(tripdata_final, member_casual, rideable_type) %>%
  rename(total = n)

hour_used_per_group <- tripdata_final %>%
  group_by(member_casual, hour_used) %>%
  summarize(total_ride = n(), avg_ride_length = mean(ride_length))

day_used_per_group <- tripdata_final %>%
  group_by(member_casual, day_used) %>%
  summarize(total_ride = n(), avg_ride_length = mean(ride_length))

month_used_per_group <- tripdata_final %>%
  group_by(member_casual, month_used) %>%
  summarize(total_ride = n(), avg_ride_length = mean(ride_length))
```
Save the tables to be used in Tableau visualization
```{r}
write_csv(member_per_group, path = "member-group-tripdata.csv")
write_csv(bike_type_per_group, path = "bike-type-group-tripdata.csv")
write_csv(hour_used_per_group, path = "hour-group-tripdata.csv")
write_csv(day_used_per_group, path = "day-group-tripdata.csv")
write_csv(month_used_per_group, path = "month-group-tripdata.csv")
```

